---
title: "Assignment 3 - ACM-S22"
output: html_notebook
author: "Linnea, Melina, Orla, and Tobias"
---

# To do
* Cleanup and comment Stan
* Model quality checks for simple Bayes (i.e., sigma)
* Model quality checks for weighted Bayes
1. prior - posterior update. this makes most sense for weighted bayes, but you still have sigma in simple bayes to check
2. prior and posterior predictive checks: running predictions from model and priors only and from models and posteriors.
3. you could check the presence of influential data points (the k pareto stuff that comes out of loo): save the loo to a variable, explore which datapoints have pareto k's > 0.7 (or 0.5 if you feel v conservative). The higher the pareto k, the harder it is for the model to predict that data point (and accordingly it changes the estimates to accommodate it better, therefore making it influential)





# Description

The purpose for this assignment is to analyze real world data using Bayesian models of cogntion.

The data comes from the social conformity experiment (<https://pubmed.ncbi.nlm.nih.gov/30700729/>), where cogsci students (in dataset 1) and schizophrenia patients + controls (dataset 2) combine their own intuition of trustworthiness of given faces to social information.

### Instructions

1.  Implement (at least) 2 models: simple Bayes vs weighted Bayes (bonus points if you also include a GLM model).
2.  Fit the models to one dataset (don't forget to explore the data first!)
3.  Check model quality
4.  Do model comparison
5.  Report (v minimal description of research question, v minimal description of data, description of models, model quality checks, report of results)

### Guide to the social conformity data

Both datasets (students and schizophrenia) include the following columns: - FaceID: an identifier of the specific face rated\
- ID: an identifier of the participant\
- Trial_Round1: in which trial the face was presented (during the first exposure)\
- Trial_Round2: in which trial the face was presented (during the second exposure)\
- FirstRating: the trustworthiness rating (1-8) given by the participant BEFORE seeing other ratings\
- OtherRating: the trustworthiness rating (1-8) given by others\
- SecondRating: the trustworthiness rating (1-8) given after seeing the others (at second exposure

**The students dataset also includes:**\
- Change: the difference between the second and the first rating\
- Class: participants belong to two different cohorts (1 and 2) tested at different times - Feedback: the difference between other rating and own first rating

**The schizophrenia dataset also includes:**\
- Group: 0 is comparison group, 1 is schizophrenia group - RT_Round1: time taken to produce the first rating of trustworthiness\
- RT_Round2: time taken to produce the second rating of trustworthiness

The schizophrenia data was collected within the study described in *Simonsen, A., Fusaroli, R., Skewes, J. C., Roepstorff, A., Mors, O., Bliksted, V., & Campbell-Meiklejohn, D. (2019). Socially learned attitude change is not reduced in medicated patients with schizophrenia. Scientific reports, 9(1), 1-11.*

*The students data is currently unpublished.*

**Data:**\
- Student data: <https://www.dropbox.com/s/r9917ta89qhwsl2/sc_students.csv?dl=0>\
- Patient data: <https://www.dropbox.com/s/td2oeos6kfrx7td/sc_schizophrenia.csv?dl=0>

# Setup

```{r Setup, message=FALSE}
library(pacman)

pacman::p_load(tidyverse, here, posterior, cmdstanr, rstan, brms)

options(mc.cores = parallel::detectCores()) # Parallelize
rstan_options(auto_write = TRUE)  # store compiled Stan models
```

```{r Function definitions}
# Scaling of the ratings for the Gaussian distribution models
scaler <- function(column, bins) {
  # We scale according to bins+2 to remove 0 and 1 from the scaling
  scaled_column <- rnorm(length(column),  column  * bins+2, 0) / 100
  return (scaled_column)
}
```

```{r Data preparation}
df <- read.csv("data/sc_schizophrenia.csv")

df <- df %>%
  # Remove NA's
  subset(complete.cases(.)) %>%
  # Setup correct column for controls (ID's above 200)
  mutate(Group = ifelse(ID > 200, 0, 1)) %>%
  select(Group, ID, FaceID, Trial_Round1, Trial_Round2,
         RT_Round1, RT_Round2, FirstRating, OtherRating, SecondRating) %>%
  mutate(FirstRating_c  = scaler(FirstRating, bins=7),
         OtherRating_c  = scaler(OtherRating, bins=7),
         SecondRating_c = scaler(SecondRating, bins=7))
```


# Explore the data

```{r Participants}
df %>%
  group_by(ID, Group) %>%
  count() %>%
  mutate(Group = ifelse(Group == 0, "Control", "Schizophrenia")) %>%
  ggplot(aes(x = Group, fill = Group)) +
    geom_histogram(stat="count", alpha=.8) +
    theme_minimal() +
    theme(legend.position="none") +
    scale_fill_brewer(palette="Dark2") +
    labs(
      title = "Number of Participants by Group",
      y     = "Count",
      x     = "",
    )
```

```{r Trials}
# Ensure that all participants have the same number of trials
df %>%
  group_by(ID, Group) %>%
  count() %>%
  ungroup() %>%
  select(n) %>%
  unique()
```

```{r}
# Plot reaction times
df %>%
  select(Group, RT_Round1, RT_Round2) %>%
  pivot_longer(cols = c(RT_Round1, RT_Round2)) %>%
  mutate(
    Group = ifelse(Group == 0, "Control", "Schizophrenia"),
    Round = ifelse(name == "RT_Round1", "Round 1", "Round 2"),
    value = value / 1000.0,
  ) %>%
  ggplot(aes(x = Group, y = value, fill = Group)) +
    geom_violin(alpha=.7) + 
    geom_point(
      shape=21, stroke=0, size=2,
      position=position_jitterdodge(),alpha=.1) +
    theme_minimal() +
    theme(legend.position="none") +
    scale_fill_brewer(palette="Dark2") +
    labs(
      title = "Reaction Times",
      y     = "Reaction Time (seconds)",
      x     = "",
    ) +
    facet_wrap(~ Round)
```

```{r Ratings}
# Plot ratings
df %>%
  select(Group, FirstRating, SecondRating, OtherRating) %>%
  pivot_longer(cols = c(FirstRating, SecondRating, OtherRating)) %>%
  mutate(
    Group = ifelse(Group == 0, "Control", "Schizophrenia"),
    Round = ifelse(
      name == "FirstRating", "First Rating",
      ifelse(name == "SecondRating", "Second Rating", "Other Rating")),
  ) %>%
  ggplot(aes(x = value, fill = Group)) +
    geom_histogram(stat="count", position="dodge", alpha=.8) +
    theme_minimal() +
    scale_fill_brewer(palette="Dark2") +
    scale_x_continuous(breaks=seq(1, 8), minor_breaks = seq(1, 8)) +
    labs(
      title = "Stimuli Rating by Round",
      y     = "Count",
      x     = "Rating",
    ) +
    facet_wrap(~ Round)

```


# Simple Bayes

```{r Stan}
data <- df %>%
  subset(Group == 1 & ID <= 103)

trials = length(data$ID)
agents = length(unique(data$ID))

stan_model <- cmdstan_model("stan_models/simpleBayes.stan")

stan_data <- list(
  N       = trials,
  agents  = agents,
  y       = matrix(data$SecondRating_c, nrow=trials, ncol=agents),
  Source1 = matrix(data$FirstRating_c, nrow=trials, ncol=agents),
  Source2 = matrix(data$OtherRating_c, nrow=trials, ncol=agents)
)

simpleBayes <- stan_model$sample(
  data = stan_data,
  seed = 123,
  chains = 1,
  parallel_chains = 1,
  iter_warmup = 1000,
  iter_sampling = 2000,
  refresh = 100
)

#saveRDS(simpleBayes, file = "stan_models/simpleBayes.rds")
#simpleBayes <- readRDS(file = "stan_models/simpleBayes.rds")

rm(trials, agents, data, stan_model, stan_data)

# Print the results
simpleBayes$summary()
```

```{r Sigma quality plot}
#  Prior- posterior update plot
posterior_draws <- as_draws_df(simpleBayes$draws(variables=c("sigma")))
draws <- as_draws_df(simpleBayes$draws(variables=c("sigma_prior"))) %>%
  mutate(sigma_posterior = posterior_draws$sigma)

draws %>%
  ggplot() +
    geom_density(aes(sigma_prior, fill="Prior"), color=NA, alpha=0.6) + 
    geom_density(aes(sigma_posterior, fill="Posterior"), color=NA, alpha=0.6) + 
    scale_fill_manual(name = "", values=c("FireBrick3", "dodgerblue3")) +
    theme_minimal() +
    labs(
      title = "Simple Bayes: Sigma Prior and Posterior",
      x     = "Sigma",
      y     = "Posterior Density"
    )

rm(posterior_draws)


```


```{r}
### THIS DOESNT WORK RIGHT NOW ###

# creating a dataframe for model predictions
prediction_df <- samples$summary() %>%
  subset(grepl("prediction", variable)) %>%
  mutate(prediction = mean / 9 * 100,
         true_rating = df_subset$SecondRating,
         trial = seq(length(variable))) %>%
  select(trial, true_rating, prediction)


# plotting the model predictions against true ratings
prediction_df %>%
  pivot_longer(cols=c("true_rating", "prediction")) %>%
  subset(trial < 30) %>%
  ggplot(aes(x = trial, y = value, color = name)) +
  geom_line()

# investigate error of model predictions
prediction_df <- prediction_df %>%
  mutate(error = abs(true_rating - prediction))

# print the summary stats
summary(prediction_df$error)
sd(prediction_df$error)

# inspect the loo
samples$loo()

```


# Weighted Bayes

```{r}
data <- df %>%
  subset(Group == 1 & ID <= 103)

trials = length(data$ID)
agents = length(unique(data$ID))

stan_model <- cmdstan_model("stan_models/weightedBayes_multilevel.stan")

stan_data <- list(
  trials  = trials,
  agents  = agents,
  choice  = matrix(data$SecondRating_c, nrow=trials, ncol=agents),
  source1 = matrix(data$FirstRating_c, nrow=trials, ncol=agents),
  source2 = matrix(data$OtherRating_c, nrow=trials, ncol=agents)
)

weightedBayes <- stan_model$sample(
  data = stan_data,
  fixed_param = FALSE, 
  seed = 123,
  chains = 1,
  parallel_chains = 1,
  iter_warmup = 1000,
  iter_sampling = 2000,
  refresh = 500
)

rm(trials, agents, data, stan_model, stan_data)

# Print the results
weightedBayes$summary()
```




